---
title: "Ant Tropic Levels from Isotope Analysis (Puerto Rico)"
author: 
  - Matthew Gilbert
  - Sofia Sarak
  - Lily Rudofksy
format: pdf
---

## Background

The Earth & Environmental Sciences (E&ES) Department at Wesleyan University offers a Capstone Project option for students in their final year. In 2025, E&ES seniors traveled to Puerto Rico to collect data for four different projects investigating various environmental themes: FIND. 

The course gives students the opportunity to brainstorm, plan, execute, and present a research project from start to finish, providing guidance and resources along the way. The seminar typically involves a travel component to collect field data, graciously funded by alumn FIND.

# Ants, Plants, and Isotopes: Investigating the Trophic Levels of Guanica Dry Forest

Our project ___

### Guanica

### Isotopes

### Research Questions

## Statistical Analyses

```{r}
#| message: false
library(here)
library(tidyverse)
library(janitor)
```

```{r}
#| message: false
ants <- read_csv(here("data/ANTSREAL.csv")) %>% 
  clean_names() %>% 
  rename(d15n = d15n_permil_vs_air,
         d13n = d13c_permil_vs_vpdb)
```


## Figures

```{r}
library(ggplot2)
```

```{r}
trail_colors <-  c("La Hoya" = '#03bb38ff', "Ojo de Agua" = '#629dffff', "Ballena" = '#f8766cff')
```


```{r}
ggplot(ants, aes(x=trail_id, y=d15n, fill=trail_id)) + 
  geom_boxplot() + 
  ylim(0,10.5) +  
  labs(title = " ",
       x=" ", 
       y=expression(delta*"15N"),
       fill = "Trail ID") +
  scale_fill_manual(values = trail_colors) +
  theme(panel.background=element_rect(fill='#d77e4850'))
```

```{r}
ggplot(ants, aes(x=d13n, y=d15n)) + geom_point() + theme_linedraw() + 
  labs(y= "delta15N permil", x="delta13C permil", title="Variation Between Ant Individuals (delta15N and delta13C)", subtitle="r^2=0.04, LOBF: y=3.15146-0.1336x") +
  geom_smooth(method="lm", se=FALSE, color="red")
```

